---
title: "Project proposal"
author: "MEG"
output: github_document
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
library(readr)
library(ggridges)
library(ggplot2)
```

## 1. Introduction


## 2. Data

```{r analysis}
victims_incidents <- read_csv("../data/victims_incidents.csv")
glimpse(victims_incidents)

```

## 3. Data analysis plan

In a choropleth map with bar chart we are hoping to visualize these variables:

Density of incidents
Date (Year)
Gender
County

In a bar chart we are hoping to visualize these variables:
Perpetrator type
Conviction


We were considering researching laws in different areas, as they may be significant in swaying the outcome to 

The preliminary data analysis shows there are variations in the amount of victims based on gender throughout time. It also shows that there is variation in the number of incidents in each state throughout time. 
```{r}
names(victims_incidents)
dim(victims_incidents)

ggplot(victims_incidents, aes(
  sex))+
  geom_bar()

victims_incidents<-victims_incidents|>
  drop_na(sex)
ggplot(victims_incidents, aes(
  incident_year_range_end, fill=sex))+
  geom_bar()

ggplot(victims_incidents, aes(
  incident_year_range_end, state, fill=state))+
  geom_density_ridges()
```


## 4. Data Ethics Review

The data ethics review section will be introduced in a separate class and is not part of the original proposal deadline.

