---
title: "Analysis of Racial Violence in the American South"
author: "MEG Team: Esme McDonald, Gus Rivers-McHugh, Maxie Langenberg"
output: github_document
---

This document should contain a detailed account of the data clean up for your data and the design choices you are making for your plots. For instance you will want to document choices you've made that were intentional for your graphic, e.g. color you've chosen for the plot. Think of this document as a code script someone can follow to reproduce the data cleaning steps and graphics in your handout.

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
library(tidyverse)
library(readr)
library(ggridges)
library(ggplot2)
library(usmap)
library(dplyr)
library(tigris)
library(stringr)
library(leaflet)
library(sf)
library(visdat)
library(naniar)
library(UpSetR)
library(devtools)
#devtools::install_github("hrbrmstr/streamgraph")
library(RColorBrewer)
library(viridis)
```

## Data Clean Up Steps for Overall Data

### Step 1: Uploading main dataset and FIPS geographic data
```{r upload data}
victims_incidents <- read_csv("../data/victims_incidents.csv")

fips_data <- tigris::counties(state = "Alabama", class = "sf") %>%
  select(fips = GEOID, county = NAME) #load fips data
``` 

### Step 2: Adding a "perp_group" column based on the four individual true/false columns for perp type. 

```{r merge-perp-groups}
victims_incidents <- victims_incidents %>%
  mutate(perp_group = case_when(
    perp_group_police == TRUE ~ "police",
    perp_group_posse == TRUE ~ "posse",
    perp_group_hate == TRUE ~ "hate",
    perp_group_mob == TRUE ~ "mob",
    TRUE ~ NA_character_
  ))
```

### Step 3: Create a count of incident total by each state 
``` {r incidents_state_map}
incident_count <- victims_incidents %>%
  group_by(state) %>%
  summarize(incident_total = n())
```



## Plots

#### Final Plot 1: Death count by perpetrator group and year

```{r year_perp_group_bar}

victims_incidents <- victims_incidents %>% 
  drop_na(perp_group)

victims_incidents$perp_group <- factor(victims_incidents$perp_group, levels = c("posse", "mob", "hate", "police"))

red_palette <- viridis(4, option = "F", begin = 0.2, end = 0.9)
red_palette <- rev(red_palette)

ggplot(victims_incidents, aes(
  incident_year_range_end, fill = perp_group)) +
  geom_bar() +
  labs(
    title = "Deaths by racial violence by year and perpetrator type",
    subtitle = "Of reported incidents in the American South between the years 1930 and 1954",
    y = "Death Count",
    x = "Year"
  ) +
  scale_fill_manual(
    name = "Perpetrator Group",
    labels = c("Posse", "Mob", "Hate Group", "Police"),
    values = red_palette
  ) +
  theme_minimal()+
  theme(
    panel.background = element_rect(fill = "#ebebeb"), 
    legend.position = "right") 

#to save image
ggsave("BARyear_perp_final.png", width = 8, height = 5)
```


### Plot 2: Anti-Black Deaths by State
```{r spaghetti plot}

data <- victims_incidents %>%
  filter(state %in% c("Alabama", "Arkansas", "Louisiana", "Mississippi", "North Carolina", "South Carolina", "Tennessee", "Texas", "Virginia")) |>
  mutate(year = incident_year_range_beg) |>
  group_by(year, state, sex) |>
  summarize(n = n())

data %>%
  drop_na(sex) |>
  ggplot( aes(x=year, y=n, group=state, color=state)) +
    geom_line(alpha = .9) +
    scale_color_viridis_d() +
    geom_vline(xintercept = 1945, linetype = "dotted", color = "red") +
  geom_vline(xintercept = 1951, linetype = "dotted", color = "red") +
  annotate("rect", 
  xmin = 1945, 
  xmax = 1951, 
  ymin = 0, 
  ymax = 20, 
  alpha = 0.2, 
  fill = "red") +
  geom_text(x = 1943, y = 20, label = "Birmingham Bombings", color = "purple", show.legend  = FALSE) +
 labs( x = "Year", y = "Death Count") +
   theme(
      plot.title = element_text(size=10)) +
     ggtitle("Anti-Black Deaths by State from 1930-1954") +
      theme_ipsum()


```

### Plot 3: Leaflet plot

```{r data wrangling}
victims_incidents <- read_csv("../data/victims_incidents.csv") # Read dataset

victims_incidents %>% 
  group_by(state) %>%
  summarise(total_incidents = n()) %>%
  arrange(desc(total_incidents)) %>% 
  print() # Count number of rows by state

all_counties_map <- victims_incidents %>% 
  mutate(county = str_replace(county, " County", "")) %>%  # because county names don't match fips data
  group_by(state, county) %>% # Select counties as a group with the state
  summarise(incident_count = n(), .groups = 'drop') %>% # Count incidents per county
  arrange(state, county) # Arrange by state and then county

counties_data <- tigris::counties(state = c("AL", "AR", "FL", "GA", "LA", "MS", "NC", "SC", "TN", "TX", "VA"), class = "sf")|> #Grab multipolygon shapefiles of each county
  arrange(STATEFP, NAME)|> # Arrange by state and then county
  rename(county = NAME) # Rename column to "county"
  
counties_data<-counties_data|> # Convert STATEFP number to state name to allow join
  mutate(state = case_when(
    STATEFP == "01" ~ "Alabama",
    STATEFP == "05" ~ "Arkansas",
    STATEFP == "12" ~ "Florida",
    STATEFP == "13" ~ "Georgia",
    STATEFP == "22" ~ "Louisiana",
    STATEFP == "28" ~ "Mississippi",
    STATEFP == "37" ~ "North Carolina",
    STATEFP == "45" ~ "South Carolina",
    STATEFP == "47" ~ "Tennessee",
    STATEFP == "48" ~ "Texas",
    STATEFP == "51" ~ "Virginia"))|>
  select(county, state, geometry)

all_counties_map<-all_counties_map|>
  mutate(county = str_replace(county, " Parish", ""))|> # Fix Louisiana
   mutate(county = str_replace(county, "Saint ", "St. ")) #Fix "Saint"

counties_data <- counties_data|>
  left_join(all_counties_map, by = c("state", "county")) # Joining datasets by state and then county. This includes counties with no incidents.

problems<- all_counties_map |>
  anti_join(counties_data, join_by(county)) # Show data that is problematic

sf_counties <- counties_data|> # Create as shapefile
  st_as_sf()

counties_data<-counties_data|>
  mutate(count_cut = cut(incident_count, breaks = c(-Inf,0.9,1.9,2.9,3.9,4.9,5.9,10.9,40.9,Inf), labels = c("Not reported","1","2","3","4","5","6-10","11-40",">40")))|>
  mutate(count_cut = ifelse(is.na(count_cut), "Not reported", as.character(count_cut))) # Create column with grouped numbers

counties_data$count_cut <- factor(counties_data$count_cut, 
                                  levels = c("Not reported", "1", "2", "3", "4", "5", "6-10", "11-40", ">40")) #Treat new column as factor
```
```{r leaflet plot} 
palcounty<-colorFactor("Reds", domain = counties_data$count_cut) #Create palette

leaflet(counties_data) |>
  addProviderTiles(providers$Jawg.Streets) |>
  addPolygons(
    fillColor = ~palcounty(counties_data$count_cut),  # Color by incident count
    weight = 0.5,
    highlightOptions = highlightOptions(weight = 1,
        color = "white",
        fillOpacity = 1,
        bringToFront = TRUE),
    opacity = 1,
    color = "black",
    fillOpacity = 1,
    popup = ~paste("State: ", state, "<br>", "County: ", county, "<br>", "Deaths reported: ", incident_count, "<br>"))|> # Popups
   addLegend(position = "topleft",
            pal = palcounty,
            values = ~counties_data$count_cut, 
            title = "Deaths reported" ,
            opacity = 1)
```
This leaflet plot we chose to have opaque county polygons because the labels bleeding through made it harder to read. We included a mouseover highlight to show clearly that the map is interactive. Clicking on each county brings up a popup menu so that more specific counts can be read, in addition to the state and county name. Grouping the death counts was chosen because the color scale was not very useful if a linear color association was chosen. Most of the data is below 5, so 1-5 are their own colors. 


### Plot 4: ___________ 


